Stage-Splitting a Modal Language

### 2014-06-24 paper/splitting.tex
Few more notes on splitting.

It's not clear whether the {\tt x} binding should be thought of as happening at stage \bbone\ or stage \bbtwo.
The variable itself is of course at stage \bbone, but the expression that is bound to it has some stage \bbtwo\ content, namely \verb|1+2|.
The solution is to think of {\tt x} as existing at both stages, and so it appears in the precomputation {\em and} the residual.
Here's the precomputation:
\begin{lstlisting} 
let x = ((), 3+4) in 
#2 x
\end{lstlisting}
and the residual:
\begin{lstlisting} 
fn l =>
  let x = (1+2, ()) in 
  (#1 x) * (#1 x) * l
\end{lstlisting}
What we see is that {\tt x}, which has type $\rmint \times \fut \rmint$ in the original program, 
has types $\rmint \times \rmint$ in the precomputation and $\rmunit \times \rmint$ in the residual.

@@

\TODO functions

@@

>>>>

It's not clear whether the {\tt x} binding should be thought of as happening at stage \bbone\ or stage \bbtwo.
The variable itself is of course at stage \bbone, but the expression that is bound to it has some stage \bbtwo\ content, namely \verb|1+2|.
The solution is to think of {\tt x} as existing at both stages, and so it appears in the precomputation {\em and} the residual.
Here's the precomputation:
\begin{lstlisting} 
let x = ((), 3+4) in 
#2 x
\end{lstlisting}
and the residual:
\begin{lstlisting} 
fn l =>
  let x = (1+2, ()) in 
  (#1 x) * (#1 x) * l
\end{lstlisting}
What we see is that {\tt x}, which has type $A = \rmint \times \fut \rmint$ in the original program, 
has types $A_1 = \rmint \times \rmint$ in the precomputation and $A_2 = \rmunit \times \rmint$ in the residual.
We call $A_1$ and $A_2$ the first and second {\em images} of $A$.

@@

The trick when splitting functions is that the contents themselves may be multi-stage.
Thus the boundary type appears as output of the first stage and input of the second stage.
This makes it hard to type the output, because the boundary types are not represented in the original function type.

@@

We express residuals as functions in the informal development above.
For reasons of concision, we equivalently represent residuals as terms open on a variable in the formal judgements.

The $\splitonesym$ judgment takes an open stage \bbone\ term to a combined term and a residual.
The $\splittwosym$ judgment takes an open stage \bbtwo\ term to a precomputation and a residual.
